### 2013-09-03 letters/neuroimage.tex
Crap draft of the neuroimage letter

\documentclass{letter}
\usepackage{hyperref}
\signature{Jon Pipitone \\
Kimel Family Translational Imaging-Genetics Laboratory \\
Research Imaging Centre \\
Centre for Addiction and Mental Health \\
Toronto, Ontario}
\address{}
\begin{document}

\begin{letter}{Editorial Board, Neuroimage}
\opening{Dear Sir or Madam:}

Please find enclosed our manuscript entitled, {\em Bootstrapping Multi-atlas
Segmentation Using Multiple Automatically Generated Templates for the
Segmentation of the Whole Hippocampus and Subfields} by Jon Pipitone and
colleagues.

In this manuscript we describe a novel automated MRI hippocampal segmentation
algorithm optimised to perform well with only a small number of manually
segmented training images. We believe this is an important contribution since
the expertise and effort needed to perform manual segmentations can be
prohibitive for many clinicians and researchers, and yet when one-of-a-kind or
modifications to existing segmentations need to be made, existing automated
methods require 30-80 training inputs.

Our automated segmentation algorithm, mischievously named MAGeT-Brain, takes
advantage of the neuroanatomical variability that exists in a population that
is being studied in order to automatically identify the hippocampus and its
subfields. Our results demonstrate consistent accuracy in the identification of
the whole hippocampus when compared to "gold-standard" manual segmentations.
Finally, we have made the our algorithm available publically online for use by
other groups, and are pursuing the the contribution of our segmentations to the
Alzheimer's Disease Neuroimaging Initiative image database.

We believe that the techniques we have developed and our findings are a
significant contribution to the neuroimaging community, specifically for those
researchers interested in large scale studies of the hippocampus in the context
of normal brain function and different forms of brain dysfunction.

We hope you find the enclosed manuscript meets the high standards of
NeuroImage.

\closing{Sincerely,}

%\ps{P.S. You can find the full text of GFDL license at}
\encl{Manuscript}

\end{letter}
\end{document}